Demand Forecasting of individual Probability Density Functions with Machine Learning

### 2020-08-05 LaTex/paper.tex
clearer evaluation section, first paragraphs of example section

\begin{abstract}
Demand forecasting is a central component for many aspects of supply chain operations, as it provides crucial input for subsequent decision making like ordering processes. While machine learning methods can significantly improve prediction accuracy over traditional time series forecasting, the calculated predictions are often mere point estimations for the conditional mean of the underlying random distribution, and the most powerful approaches, like deep learning, are usually opaque in terms of how its individual predictions can be interpreted. Using the novel supervised machine learning method ``Cyclic Boosting'', complete individual probability density functions in form of negative binomial distributions can be predicted instead of simple point estimates. Additionally, each single prediction is explainable which is a major benefit in practice as individual forecasts can be understood by the practitioner and ``black-box'' models can be avoided. Also, we describe new techniques for both qualitative and quantitative evaluation of predicted probability density functions by comparison with observed data.
\end{abstract}

{Keywords: \textbf{machine learning, demand forecasting}}

PDF
\section{Introduction}

Demand forecasting is one of the main challenges for retailers and at the core of business operations. Due to its stochastic nature, demand is difficult to forecast as it depends on many influencing factors and the realized demand can be interpreted as a random variable that is described by an appropriate probability density function (PDF). In order to make operational decisions, an optimal point estimator has to be defined, that can be used to derive ordering decisions used in the replenishment process of the retailer. Demand estimation is further complicated by the fact that retailers typically only observe realized sales rather than the actual demand, and in case the demand exceeds the current stock level the data become censored. The ordering decision process is complicated by a range of factors: Even in the case of accurate  demand forecasts, the decision maker has to balance conflicting metrics to reach an optimal decision: Ordering to few items may result in stockout situations leading to unrealized demand and unsatisfied customers. Ordering too many items results in excess inventory which increases transport and storage costs and, in the case of perishable goods, excessive waste, as spoilt items need to be disposed of at additional cost and potentially even environmental impact. This situation is particularly noticeable in the so called "ultra-fresh'' category, which includes items such as bakery products, ready-meals, fresh diary products, or certain meat products such as ground meat. These items typically have a shelf-life of less than a business day to a few business days at most, with a continuous spectrum depending on the exact item. In many situations, additional constraints have to be considered to reach an optimal ordering decision: Delivery cycles of items may vary depending on the type of item and the wholesaler or manufacturer from which they are procured. Retailers also operate at a given service level to guarantee that a certain level of demand can be fulfilled. The exact service level typically depends on the overall business policy of the retailer and may also depend on individual products, ranging from "never-out-of-stock'' items to a service level exceeding e.g. 90\%.

The availability of Big Data allows capturing, storing and processing a vast amount of data associated with demand, such as historic sales records, information about promotional events or advertisements, pricing information, local weather at retail locations, seasonal information as well as a wide range of further variables. Modern machine learning algorithms can then be used to predict the per-item demand distribution, corrected for censored data, from which an optimal point estimator can be derived to be used in the subsequent ordering decision. It is important to note that demand as a random variable is not identically and independently distributed (i.i.d.). While the probability distribution describing the demand can be attributed to a given family or parametrization, the exact parameters vary: Seasonal effects, finite life cycles of products and the introduction of new products influence the demand distribution, as well as the local weather at the retail location or the retail location itself in terms of size, assortment range, customer diversity and other factors. The retailers themselves also actively influence demand by using advertisements to highlight products, offering rebates or discounts for specific products as well as pursuing an active pricing strategy. This means that while we can generally assume that demand follows a specific type of probability distribution, its parameters are unique to the  instance for which an estimate is required. For example, the probability distribution governing the demand of a particular item is specific to the item, date and retail location for which the forecast is made and depends on a wide range of further influencing factors.

@@

Inventory management offers a rich theory and the extensive body of research can be broadly grouped into two categories, where the inventory control problem is either based on some knowledge of the underlying demand  distribution or an integrated approach that seeks to map directly from the available data (historic sales records and further variables) to the ordering decision. This approach is often referred to as "data-driven newsvendor'' and discussed  e.g. in \cite{beutel2012safety,ban2019big,bertsimas2020predictive, oroojlooyjadid2020applying}. It aims to avoid estimating the underlying probability distribution and use the available data to derive the operational decisions directly.  An overview of a range approaches can also be found in \cite{huber2019data}.

Although this approach seems preferable at first glance, since it avoids determining the full demand distribution and results directly in the desired operational decision (the order quantity), it faces several drawbacks. First of all, the full probability distribution for the demand of a specific item at a given sales location and business day includes all available information including the uncertainty of the modelled demand. This can be used to simulate the performance on a per-item level and e.g. optimize the impact on business strategy decisions on conflicting metrics such as stock out- and waste-rate. By forecasting the full demand distribution as opposed to point estimators such as the expected demand or the median of the distribution, the forecast quality can be evaluated for quantiles of the distribution, including the often extensive tails of the distribution. Additionally, a purely data-driven approach going from the observed data directly to the operational decision (such as the order quantity) does not allow to analyze the data-generating process, i.e. the mechanism behind the stochastic behavior of the customer demand. However, modelling demand directly is vital if a causal analysis is planned at a later stage or independently, for example to study the effect promotions, advertisements,  price changes or other influencing factors in either Pearl's do-calculus \cite{PearlCausality} or Rubin's potential outcomes framework \cite{rubin1974estimating}. Using an operational quantity such as the order quantity will in most cases be an insufficient proxy of the quantity of interest (customer demand) and likely lead to unnecessary causal pathways that may not be able to be fully controlled for. From a practical perspective, separating the demand forecast from the operational decisions (i.e. calculating the  order quantities for the next delivery cycle) also allows to evaluate longer-term planning and reduces the complexity as it avoids coupling the complex delivery schedules of multiple wholesalers and manufacturers with the forecast of customer demand. This also allows to share long-term demand predictions with other business units or external vendors  and wholesalers to ease their planning for the production and supply-chain processes upstream of the retailer. From the perspective of industrial practice, modelling the demand separately from deriving the subsequent orders  has the additional benefit that multiple retail chains can benefit from any improvement in the model description even if the concrete retailers are unrelated to each other. For example, if a particular effect was identified at some retailer A and included in the machine learning model, the improvement can be rolled out immediately or on request to all other retailers using this system on a planetary scale without having to adapt the underlying machine learning model for each retailer individually.

@@

In the following, we will demonstrate how the explainable machine learning algorithm Cyclic Boosting \cite{Wick2019} can be used to model the demand distribution at the granularity needed by the retailer. Typically this means that the full demand distribution has to be estimated per SKU for each sales location and opening day, conditional on a wide range of variables such as weather, prices, promotions, etc. In contrast to using a "black-box'' machine learning model, using Cyclic Boosting allows to interpret how each individual prediction was made and what the most important variables are per individual prediction.

Additionally, we show how the cumulative distribution function (CDF) or inverse quantiles can be used to accurately asses the forecast quality of the full predicted demand distribution, including the tails of the distribution. This allows to verify that the predicted demand distribution accurately reflects the observed data and can hence be used both to derive operational decisions such as  order quantities as well as strategic business decisions by the retailer.

@@

To predict an individual PDF by means of a parametric approach, one has to rely on a model assumption about the underlying distribution of the random variable to be predicted. In the following, we describe such a parametric approach with the model assumption of a negative binomial probability distribution. For a PDF prediction under a negative binomial model assumption, there is the need for the estimation of two parameters, for example its mean and its variance.

This can be done using two independent models, one to estimate the mean and the other for the variance. At least in principle, any method can be used. However, as argued above, in the case of demand forecasting, each prediction is highly specific to the circumstances in which it is used (such as product ID, date, and store location) and may depend on  a multitude of describing variables (features). Machine learning algorithms are ideally suited for this task and in the following we will use the Cyclic Boosting algorithm.

This means we use two subsequent, independent Cyclic Boosting models in order to estimate individual PDFs, the first to estimate the mean and the second to estimate the variance. The features may or may not differ between the mean and variance estimation models, and one might even include the corresponding predicted mean from the mean model as feature in the variance model. The assigned mean and variance predictions can then be used to generate individual PDFs according to a functional negative binomial distribution assumption for each sample. While any machine learning algorithm outputting a conditional mean as prediction may be used to get an approximate estimate for the mean of an underlying negative binomial distribution, it is preferable to choose an algorithm using a negative binomial assumption for its inner regularization methods, like Cyclic Boosting in multiplicative regression mode.

In the following, after a brief recap of the fundamental ideas of Cyclic Boosting, we describe a method to predict the variance of a negative binomial PDF along with its mean individually for each sample by means of Cyclic Boosting.

\subsection{Cyclic Boosting Algorithm}
\label{sec:CB}

Cyclic Boosting \cite{Wick2019} is a novel machine learning algorithm that can be categorized as a generalized additive model with a cyclic coordinate descent optimization featuring a boosting-like update of parameters. The major benefits of Cyclic Boosting are that it is not only extremely performant in practical applications, like demand forecasting, at large scale, but also fully explainable. Traditional machine learning algorithms are typically "black box'' approaches, where the individual decision cannot be explained. Cyclic Boosting on the other hand allows to follow how each individual prediction was made. The main idea of this algorithm is the following: When used for regression, denoted as multiplicative regression mode, the predictions $\hat{y_i}$ of the  target variable $Y \in [0,\infty)$ can be calculated in the following way:

\begin{equation} \label{eqn:cb}
\hat{y}_i = \mu \cdot \prod \limits_{j=1}^p f^k_j \quad \text{with}\; k=\{ x_{j,i} \in b^k_j\}
\end{equation}

The values $f^k_j$ are the model parameters that are determined from features $j$. Each feature is discretized appropriately into $k$ bins to reflect the specific behaviour of the feature. The global mean $\mu$ is determined from all target values $y$ observed in the data. The factors $f^k_j$ are determined iteratively until convergence is reached and regularization techniques are applied to avoid overfitting and improve the generalization ability of the algorithm. The deviations of the different factors from $f^k_j=1$ are then directly interpretable in terms of how much a specific feature contributes to each individual prediction.

The following meta-algorithm is used to determine the model parameters $f^k_j$ from the training data:
\begin{enumerate}
\item{Calculate the global average $\mu$ from all observed $y$ across all bins $k$ and
features $j$.}
\item{Initialize the factors $f^k_j \leftarrow 1$}
\item{Cyclically iterate through features $j = 1,...,p $ and calculate in turn for each bin $k$ the partial factors $g$ and corresponding aggregated factors $f$, where indices $t$ (current iteration) and $\tau$ (current or preceding iteration) refer to iterations of full feature cycles as the training of the algorithm progresses:
\begin{equation} \label{factors}
g^k_{j,t} = \frac{\sum \limits_{x_{j,i} \in b^k_j} y_i}{\sum \limits_{x_{j,i} \in b^k_j} \hat{y}_{i,\tau}}
\;\; \mathrm{where} \; \; f^k_{j,t} = \prod \limits_{s=1}^t g^k_{j,s}
\end{equation}
This means $g$ is the factor that is multiplied to $f_{t-1}$ in each iteration. Here, $\hat{y}_\tau$ is calculated according to eqn. \ref{eqn:cb} with the current values of the aggregated factors $f$:
\begin{equation} \label{factors3}
\hat{y}_{i,\tau} = \mu \cdot \prod \limits_{j=1}^p f^k_{j,\tau}
\end{equation}
To be precise, the determination of $g^k_{j,t}$ for a specific feature $j$ employs $f^k_{j,t-1}$ in the calculation of $\hat{y}$. For the factors of all other features, the newest available values are used, i.e., depending on the sequence of features in the algorithm, either from the current ($\tau=t$) or the preceding iteration ($\tau=t-1$).
}
\item{Quit when stopping criteria, e.g. the maximum number of iterations or no further improvement of an error metric such as the mean absolute deviation (MAD) or mean squared error (MSE), are met at the end of a full feature cycle.}
\end{enumerate}

\subsection{Variance Estimation by Cyclic Boosting in Negative Binomial Width Mode}

Cyclic Boosting in a modified form of the one briefly described in sec. \ref{sec:CB}, namely the negative binomial width mode presented in the following, can be used for the second machine learning model to predict the negative binomial variance associated with the mean predicted by the first model.

For regression of a target variable $y$ taking the values $y = 0, 1, 2, ...$, the parametrization of the negative binomial mass function by means of mean $\mu$ and dispersion parameter $r$ can be specified as \cite{hilbe2011negative}:

\begin{equation} \label{eqn:nbinom}
\mathrm{NB}(y; \mu, r) = \frac{\Gamma(r + y)}{y! \cdot \Gamma(r)} \cdot \left(\frac{r}{r + \mu}\right)^r \cdot \left(\frac{\mu}{r + \mu}\right)^y,
\end{equation}

Bounding the inverse of the dispersion parameter $1/r$ to the interval $[0, 1]$ (corresponding to bounding $r$ to the interval $[1, \infty]$), the variance $\sigma^2$ can be calculated from $\mu$ and $r$ via:

\begin{equation} \label{eqn:variance_r}
\sigma^2 = \mu + \frac{\mu^2}{r}
\end{equation}

An estimation of the dispersion parameter $\hat{r}$ can be achieved by minimizing the loss function defined in \eqn \eqref{eqn:loss_likelihood}, expressed as negative log-likelihood function of a negative binomial distribution, with respect to the Cyclic Boosting parameters $f^k_j$ constituting the model of $\hat{r_i}$ according to \eqn \eqref{eqn:r}, over all input samples $i$, where the mean values $\hat{\mu_i}$ are fixed to the corresponding predictions from the first machine learning model outputting the conditional mean.

@@

So, the values $\hat{r_i}$ are estimated by means of the Cyclic Boosting model parameters $f^k_j$ for each feature $j$ and bin $k$. For any concrete observation $i$, the index $k$ of the bin is determined by the observation of $x_{j,i}$ and the subsequent look-up into which bin this observation falls. Just like in sec. \ref{sec:CB}, the model parameters $f^k_j$ correspond to factors with values in $[0, \infty]$ and neutral factor $1$. Analog to the classification mode of Cyclic Boosting described in \cite{Wick2019}, the structure of \eqn \eqref{eqn:r} can be interpreted as inverse of a logit link function.

As described in \cite{Wick2019}, the parameter estimation in Cyclic Boosting is an iterative method corresponding to a cyclic coordinate descent, processing one feature with all its bins at a time until convergence. Unlike in the basic multiplicative regression mode of Cyclic Boosting described in sec. \ref{sec:CB}, the minimization of the loss function in \eqn \eqref{eqn:loss_likelihood} cannot be solved analytically and requires a numerical method, for example a random search. All other advantages of Cyclic Boosting, like for example individual explainability of predictions, remain valid for its negative binomial width mode.

Finally, the variance $\hat{\sigma}^2_i$ can be estimated from the dispersion parameter $\hat{r_i}$ by means of \eqn \eqref{eqn:variance_r}. And so, together with the estimated mean parameter $\hat{\mu_i}$ from the previous mean estimation model, all parameters of the negative binomial distribution NB for the prediction of an individual observation $i$ are given.

@@

In the following, we will show how histograms of the CDF and inverse quantiles can be used to determine the accuracy of the predicted PDFs. Although we limit the following discussion to the negative binomial model, the method can be applied generally to any representation of a probability density distribution, even if the PDF is obtained empirically.

@@

As a first step, we use the inverse quantile function or probability integral transform (PIT), see e.g. \cite{Angus1994,casella2002statistical}. This offers a first qualitative assessment by looking at the distribution of the quantiles of the predicted density via the CDF. In case of a correctly calibrated predicted PDF, this distribution is expected to be uniform and any deviation can be interpreted as a hint that the predicted probability function is not correct \cite{diebold1998vevaluating}.

\subsubsection{Histogram of the CDF}
\label{sec:cdf_histo}

@@

If the CDF is continuous and strictly increasing, then the inverse of the CDF $F^{-1}(y)$ exists and is a unique real-valued number $x$ for each $y \in [0,1]$ so that we can write $F(x) = y$. The latter defines the inverse quantile function because we can define the quantile $\tau$ of the probability distribution $f(x)$ as:

@@

\begin{figure}
\begin{center}
\includegraphics[width=8cm]{../figures/PdfCdf}
\caption{\label{fig:PdfCdf} Probability distribution function and cumulative distribution funtion of a normal distrubution.}
\end{center}
\end{figure}

We can then interpret the CDF as a new variable $s = F(t)$, meaning that $F$ becomes a transformation that maps $t$ to $s$, i.e. $F:t \to s$. Accordingly,  $\lim_{t \to -\infty}s(t) = 0$ and $\lim_{t \to \infty}s(t) = 1$ and $s$ can be intuitively interpreted as the fraction of the distribution of $t$ with values smaller than $t$ from the definition of the CDF. This implies that the probability distribution of $s$, $g(s)$, is constant in the interval $s\in [0,1]$ in which $s$ is defined and $s$ can be interpreted as the cumulative distribution of its own probability distribution:

@@

In case of discrete probability functions such as the negative binomial function, the same argument still holds but the definition of the inverse quantile function is replaced by the generalized inverse: $F^{-1}(y) = \mathrm{\inf \left \{x : F(x)>y\right  \} }$ for $y \in [0,1]$, see e.g. \cite[p. 54]{casella2002statistical}. In order to obtain a uniform distribution for discrete PDFs that is comparable to the case of continuous distributions, the histogram holding the inverse quantiles or values of the CDF is filled using random numbers according to the intervals of the CDF. For example, if the sales of zero items accounts for 75\% of the observed sales distribution for this item, the value of the CDF function that is used to fill the histogram is randomly chosen in the interval $[0,0.75]$. Proceeding similarly for all other observed values, the resulting histogram of CDF values should again be uniform as in the case of a continuous PDF.

>>>>

\begin{abstract}
Demand forecasting is a central component for many aspects of supply chain operations as it provides crucial input for subsequent decision making like ordering processes. While machine learning methods can significantly improve prediction accuracy over traditional time series forecasting, the calculated predictions are often mere point estimations for the conditional mean of the underlying probability distribution, and the most powerful approaches, like deep learning, are usually opaque in terms of how its individual predictions can be interpreted. Using the novel supervised machine learning method ``Cyclic Boosting'', complete individual probability density functions in form of negative binomial distributions can be predicted instead of simple point estimates. While metrics evaluating simple point estimates are widely used, methods for assessing the accuracy of predicted distributions are rare and this work proposes new techniques for both qualitative and quantitative evaluation methods. Additionally, each single prediction obtained with this framework is explainable, which is a major benefit in practice as individual forecasts can be understood by the practitioner and ``black-box'' models can be avoided. 
\end{abstract}

{Keywords: \textbf{explainable machine learning, demand forecasting, probability distribution}}

\section{Introduction}

Demand forecasting is one of the main challenges for retailers and at the core of business operations. Due to its stochastic nature, demand is difficult to forecast as it depends on many influencing factors and the realized demand can be interpreted as a random variable that is described by an appropriate probability density function (PDF). In order to make operational decisions, an optimal point estimator has to be defined, that can be used to derive ordering decisions used in the replenishment process of the retailer. Demand estimation is further complicated by the fact that retailers typically only observe realized sales rather than the actual demand, and in case the demand exceeds the current stock level the data become censored. The ordering decision process is complicated by a range of factors: Even in the case of perfect demand forecasts, the decision maker has to consider lot-sizes defined by the wholesaler or manufacturer as well as to balance conflicting metrics to reach an optimal decision: Ordering too few items may result in stock-out situations leading to unrealized demand and unsatisfied customers. Ordering too many items results in excess inventory which increases transport and storage costs and, in the case of perishable goods, excessive waste, as spoilt items need to be disposed of at additional cost and potentially even environmental impact. This situation is particularly noticeable in the so called "ultra-fresh'' category, which includes items such as bakery products, ready-meals, fresh diary products, or certain meat products such as ground meat. These items typically have a shelf-life ranging from less than a business day to a few business days at most, with a continuous spectrum in between, depending on the exact item. In many situations, additional constraints have to be considered to reach an optimal ordering decision: Delivery cycles of items may vary depending on the type of item and the wholesaler or manufacturer from which they are procured. Retailers also operate at a given service level to guarantee that a certain level of demand can be fulfilled. The exact service level typically depends on the overall business policy of the retailer and may also depend on individual products, ranging from "never-out-of-stock'' items to a service level exceeding e.g. 90\%.

The availability of Big Data allows capturing, storing and processing a vast amount of data associated with demand, such as historic sales records, information about promotional events or advertisements, pricing information, local weather at retail locations, seasonal information as well as a wide range of further variables. Modern machine learning algorithms can then be used to predict the per-item demand distribution, corrected for censored data, from which an optimal point estimator can be derived to be used in the subsequent ordering decision. It is important to note that demand as a random variable is not identically and independently distributed (i.i.d.). While the probability distribution describing the demand can be attributed to a given family or parameterization, the exact parameters vary: Seasonal effects, finite life cycles of products and the introduction of new products influence the demand distribution, as well as the local weather at the retail location or the retail location itself in terms of size, assortment range, customer diversity and other factors. The retailers themselves also actively influence demand by using advertisements to highlight products, offering rebates or discounts for specific products as well as pursuing an active pricing strategy. This means that while we can generally assume that demand follows a specific type of probability distribution, its parameters are unique to the  instance for which an estimate is required. For example, the probability distribution governing the demand of a particular item is specific to the item, date and retail location for which the forecast is made and depends on a wide range of further influencing factors.

@@

Inventory management offers a rich theory and the extensive body of research can be broadly grouped into two categories, where the inventory control problem is either based on some knowledge of the underlying demand  distribution or an integrated approach that seeks to map directly from the available data (historic sales records and further variables) to the ordering decision. This approach is often referred to as "data-driven newsvendor'' and discussed  e.g. in \cite{beutel2012safety,ban2019big,bertsimas2020predictive, oroojlooyjadid2020applying}. It aims to avoid estimating the underlying probability distribution for demand and use the available data to derive the operational decisions (the order quantity) directly.  An overview of a range approaches can also be found in \cite{huber2019data}.

Although this approach seems preferable at first glance, since it avoids determining the full demand distribution and results directly in the desired operational decision (the order quantity), it faces several drawbacks. First, the full probability distribution for the demand of a specific item at a given sales location and business day includes all available information including the uncertainty of the modelled demand. This can be used to simulate the performance on a per-item level and e.g. optimize the impact on business strategy decisions on conflicting metrics such as stock out- and waste-rate. By forecasting the full demand distribution as opposed to point estimators such as the expected demand or the median of the distribution, the forecast quality can be evaluated for quantiles of the distribution, including the often extensive tails of the distribution. Additionally, a purely data-driven approach going from the observed data directly to the operational decision (such as the order quantity) does not allow to analyze the data-generating process, i.e. the mechanism behind the stochastic behavior of the customer demand. However, modelling demand directly is vital if a causal analysis is planned at a later stage or independently, for example to study the effect promotions, advertisements,  price changes or other influencing factors in either Pearl's do-calculus \cite{PearlCausality} or Rubin's potential outcomes framework \cite{rubin1974estimating}. Using an operational quantity such as the order quantity will in most cases act as an insufficient proxy of the quantity of interest (customer demand) and likely lead to unnecessary causal pathways that may not be able to be fully controlled for. From a practical perspective, separating the demand forecast from the operational decisions (i.e. calculating the  order quantities for the next delivery cycle) also allows to evaluate longer-term planning and reduces the complexity as it avoids coupling the complex delivery schedules of multiple wholesalers and manufacturers with the forecast of customer demand. This also allows to share long-term demand predictions with other business units or external vendors  and wholesalers to ease their planning for the production and supply-chain processes upstream of the retailer. From the perspective of industrial practice, modelling the demand separately from deriving the subsequent orders  has the additional benefit that multiple retail chains can benefit from any improvement in the model description even if the concrete retailers are unrelated to each other. For example, if a particular effect was identified at some retailer A and included in the machine learning model, the improvement can be rolled out immediately or on request to all other retailers using this system on a planetary scale without having to adapt the underlying machine learning model for each retailer individually.

@@

This work demonstrates how the explainable machine learning algorithm Cyclic Boosting \cite{Wick2019} can be used to model the demand distribution at the granularity needed by the retailer. Typically this means that the full demand distribution has to be estimated per SKU for each sales location and opening day, conditional on a wide range of variables such as weather, prices, promotions, etc. In contrast to using a "black-box'' machine learning model, using Cyclic Boosting allows to interpret how each individual prediction was made and what the most important variables are per individual prediction.

Additionally, we show how the cumulative distribution function (CDF) can be used to accurately asses the forecast quality of the full predicted demand distribution, including the tails of the distribution. This allows to verify that the predicted demand distribution accurately reflects the observed data and can hence be used both to derive operational decisions such as  order quantities as well as strategic business decisions by the retailer or gain further insights into customer behavior using for example causal modelling.

@@

To predict an individual PDF using a parametric approach, one has to rely on a model assumption about the underlying distribution of the random variable to be predicted. As discussed earlier, the negative binomial probability distribution (NBD) is well routed in theoretical arguments to model customer demand. Its parameters can be modelled by two independent models, one to estimate the mean and the other for the variance. At least in principle, any method can be used. However, in the case of demand forecasting, each prediction is highly specific to the circumstances in which it is used (such as SKU, opening day, and store location) and may depend on a multitude of describing variables or features such as sales location, weather, price, and so forth. It should be noted that these parameters are not independent between products. For example, a promotion applied to one product can, at least in principle, affect the sales of related products within the assortment. This implies that the demand forecasts cannot be treated as individual newsvendor-type predictions but need to be modelled holistically. Machine learning algorithms are ideally suited for this task and in the following we will use the Cyclic Boosting algorithm to benefit in particular from explainable decisions rather than black-box approaches. Furthermore, the regularization approach used during training of the Cyclic Boosting algorithm allows a dedicated treatment of the underlying NBD model, which is another major benefit compared to a standard ''off-the-shelf'' machine learning algorithm. This means we use two subsequent Cyclic Boosting models in order to estimate the parameters of each individual PDF that we need to forecast, where the first model is used to estimate the mean and the second to estimate the variance. The features may or may not differ between the mean and variance estimation models. It can also be beneficial to include the corresponding predicted mean from the mean model as feature in the variance model. The assigned mean and variance predictions can then be used to generate individual PDFs using the parameterization of the NBD for each sample.

In the following, after a brief recap of the fundamental ideas of Cyclic Boosting, we describe a method to predict mean and variance for individual NBD models using Cyclic Boosting.

\subsection{Cyclic Boosting Algorithm: Mean estimation}
\label{sec:CB}

Cyclic Boosting \cite{Wick2019} is a type of generalized additive models using a cyclic coordinate descent optimization and featuring a boosting-like update of parameters. Major benefits of Cyclic Boosting are its accuracy, performance even at large scale and providing fully explainable predictions which are of vital importance in practical applications, in particular for multi-national retail vendors.

The main idea of this algorithm is the following: First, each feature, denoted by index $j$, is discretized appropriately into $k$ bins to reflect the specific behaviour of the feature. The global mean $\mu$ is determined from all target values $y$ of the  target variable $Y \in [0,\infty)$ observed in the data. Single data records, for example the sales of a specific SKU along with all relevant features, are indexed by $i$.
The individual predictions $\hat{y_i}$  can then be calculated as:
\begin{equation} \label{eqn:cb}
\hat{y}_i = \mu \cdot \prod \limits_{j=1}^p f^k_j \quad \text{with}\; k=\{ x_{j,i} \in b^k_j\}
\end{equation}
The factors $f^k_j$ are the model parameters that are determined iteratively from the features, and are determined iteratively until the algorithm converges. During training,  regularization techniques are applied to avoid overfitting and improve the generalization ability of the algorithm. The deviation of each factor from $f^k_j=1$ can then be used to explain how a  specific feature contributes to each individual prediction.

In detail, the following meta-algorithm describes how the model parameters $f^k_j$ are obtained from the training data:
\begin{enumerate}
\item{Calculate the global average $\mu$ from all observed $y$ across all bins $k$ and features $j$.}
\item{Initialize the factors $f^k_j \leftarrow 1$}
\item{Cyclically iterate through features $j = 1,...,p $ and calculate in turn for each bin $k$ the partial factors $g$ and corresponding aggregated factors $f$, where indices $t$ (current iteration) and $\tau$ (current or preceding iteration) refer to iterations of full feature cycles as the training of the algorithm progresses:
\begin{equation} \label{factors}
g^k_{j,t} = \frac{\sum \limits_{x_{j,i} \in b^k_j} y_i}{\sum \limits_{x_{j,i} \in b^k_j} \hat{y}_{i,\tau}}\;\; \mathrm{where} \; \; f^k_{j,t} = \prod \limits_{s=1}^t g^k_{j,s}
\end{equation}
Here,  $g$ is a factor that is multiplied to $f_{t-1}$ in each iteration. The current prediction, $\hat{y}_\tau$, is calculated according to eqn. \ref{eqn:cb} with the current values of the aggregated factors $f$:
\begin{equation} \label{factors3}
\hat{y}_{i,\tau} = \mu \cdot \prod \limits_{j=1}^p f^k_{j,\tau}
\end{equation}
To be precise, the determination of $g^k_{j,t}$ for a specific feature $j$ uses $f^k_{j,t-1}$ in the calculation of $\hat{y}$. For the factors of all other features, the newest available values are used, i.e., depending on the sequence of features in the algorithm, either from the current ($\tau=t$) or the preceding iteration ($\tau=t-1$).}
\item{Quit when stopping criteria, e.g. the maximum number of iterations or no further improvement of an error metric such as the mean absolute deviation (MAD) or mean squared error (MSE), are met at the end of a full feature cycle.}
\end{enumerate}

\subsection{Cyclic Boosting Algorithm: Width estimation}
\label{sec:cb_width}

In the previous section, the general Cyclic Boosting algorithm was used to estimate the mean of the NBD model. In order to predict the variance of the NBD model (associated with the mean predicted before), we modify the algorithm like described in the following.

When looking at the demand of individual SKUs, the target variable $y$ has the values $y = 0, 1, 2, ...$ and the NBD model can be parameterized as in \cite{hilbe2011negative}:
\begin{equation} \label{eqn:nbinom}
\mathrm{NB}(y; \mu, r) = \frac{\Gamma(r + y)}{y! \cdot \Gamma(r)} \cdot \left(\frac{r}{r + \mu}\right)^r \cdot \left(\frac{\mu}{r + \mu}\right)^y,
\end{equation}
where $\mu$ is the mean of the distribution and $r$ a dispersion parameter.

By bounding the inverse of the dispersion parameter $1/r$ to the interval $[0, 1]$ (corresponding to bounding $r$ to the interval $[1, \infty]$), the variance $\sigma^2$ can be calculated from $\mu$ and $r$ via:
\begin{equation} \label{eqn:variance_r}
\sigma^2 = \mu + \frac{\mu^2}{r}
\end{equation}

The estimate of the dispersion parameter $\hat{r}$ can then be calculated by minimizing the loss function defined in \eqn \eqref{eqn:loss_likelihood}, which is expressed as negative log-likelihood function of a negative binomial distribution. Using Cyclic Boosting, the minimization over all input samples $i$ is performed with respect to the Cyclic Boosting parameters $f^k_j$, constituting the model of $\hat{r_i}$, according to \eqn \eqref{eqn:r}, where the estimates for the mean $\hat{\mu_i}$ are fixed to the values obtained in the previous step (described in sec. \ref{sec:CB}).

@@

In other words, the values $\hat{r_i}$ are estimated via learning the Cyclic Boosting model parameters $f^k_j$ for each feature $j$ and bin $k$ from data. For any concrete observation $i$, the index $k$ of the bin is determined by the value of the feature $x_{j,i}$ and the subsequent look-up into which bin this observation falls. Like in sec. \ref{sec:CB}, the model parameters $f^k_j$ correspond to factors with values in $[0, \infty]$ and again values deviating from $f^k_j=1$ can be used to explain the relative importance of a specific feature contributing to individual predictions. Note that the structure of \eqn \eqref{eqn:r} can be interpreted as inverse of a logit link function in the same way as explained in  \cite{Wick2019} when Cyclic Boosting is used for classification tasks.

The Cyclic Boosting algorithm is trained iteratively using cyclic coordinate descent, processing one feature with all its bins at a time until convergence is reached. Unlike in the basic multiplicative regression mode of Cyclic Boosting described in sec. \ref{sec:CB}, the minimization of the loss function in \eqn \eqref{eqn:loss_likelihood} cannot be solved analytically and has to be done numerically, for example using a random search. All other advantages of Cyclic Boosting, like for example individual explainability of predictions, remain valid for its negative binomial width mode.

Finally, the variance $\hat{\sigma}^2_i$ can be estimated from the dispersion parameter $\hat{r_i}$ using \eqn \eqref{eqn:variance_r}. Using  the individual predicted mean $\hat{\mu_i}$ from the first step, the model is fully specified for each individual prediction $i$.

@@

In the following, we will show how different visualizations of the observed cumulative distribution function (CDF) values can be used to evaluate the quality of the predicted PDFs. Although we limit the following discussion to the negative binomial model, the method can be applied generally to any representation of a probability density distribution, even if the PDF is obtained empirically.

@@

In the simplest case, we only have one model with one set of model parameters to cover all predictions. In this case, the evaluation of the full probability distribution is straight-forward: We would fill a histogram of all observed values, such as sales records, and overlay this with the single model, such as a negative binomial with predicted parameters, that is used for all observations. Then, we compare the model curve directly with the observations, using statistical tests such as the Kolmogorov-Smirnov test.

In practical applications however, we have a large number of effective prediction models, since although we always use the same model parameterization, such as the negative binomial distribution, its parameters have to be determined at the required level of granularity. For example, for daily orders, we need to predict the parameters of the negative binomial distribution for each location, sales day, and product. Unlike the simple case discussed above, where we had many observations to compare the prediction model to, we now have just a single observation per prediction, meaning that we cannot use statistical tests directly.

\subsubsection{Histogram of CDF Observations}
\label{sec:cdf_histo}

For a first qualitative assessment, we make use of the probability integral transform, see e.g. \cite{Angus1994,casella2002statistical}, which states that a random variable distributed according to the CDF of another random variable is uniformly distributed. So, in case of a correctly calibrated PDF prediction, the distribution of the corresponding actually observed CDF values of the individual PDF predictions is expected to be uniform, {\em regardless} of the shape of the predicted distribution, and any deviation can be interpreted as a hint that the predicted PDF is not fully correct \cite{diebold1998vevaluating}.

@@

If the CDF is continuous and strictly increasing, then the inverse of the CDF, $F^{-1}(y)$, exists and is a unique real-valued number $x$ for each $y \in [0,1]$, so that we can write $F(x) = y$. The latter defines the inverse quantile function, because we can define the quantile $\tau$ of the probability distribution $f(x)$ as:

@@

\begin{figure}
\begin{center}
\includegraphics[width=8cm]{../figures/PdfCdf}
\caption{\label{fig:PdfCdf} Probability distribution function and cumulative distribution function of a normal distribution.}
\end{center}
\end{figure}

We can then interpret the CDF as a new variable $s = F(t)$, meaning that $F$ becomes a transformation that maps $t$ to $s$, i.e. $F:t \to s$. Accordingly,  $\lim_{t \to -\infty}s(t) = 0$ and $\lim_{t \to \infty}s(t) = 1$ and $s$ can be intuitively interpreted as the fraction of the distribution of $t$ with values smaller than $t$ from the definition of the CDF. This implies that the probability distribution of $s$, $g(s)$, is constant in the interval $s \in [0,1]$ in which $s$ is defined, and $s$ can be interpreted as the cumulative distribution of its own probability distribution:

@@

In case of discrete probability functions, such as the negative binomial function, the same argument still holds, but the definition of the inverse quantile function is replaced by the generalized inverse: $F^{-1}(y) = \mathrm{\inf \left \{x : F(x)>y\right  \} }$ for $y \in [0,1]$, see e.g. \cite[p. 54]{casella2002statistical}. In order to obtain a uniform distribution for discrete PDFs that is comparable to the case of continuous distributions, the histogram holding the inverse quantiles or values of the CDF is filled using random numbers according to the intervals of the CDF. For example, if the sales of zero items accounts for 75\% of the observed sales distribution for this item, the value of the CDF function that is used to fill the histogram is randomly chosen in the interval $[0,0.75]$. Proceeding similarly for all other observed values, the resulting histogram of CDF values should again be uniform as in the case of a continuous PDF.
